### 2015-07-30 writeup/referee_response.txt
Change to bias answer.

We thank the review for the helpful suggestions and positive review. 
In the revised version of the manuscript, changes are highlighted in blue for
the reviewers and editor's convenience. Our responses to each of the reviewers
comments are in-line below.

> ----------------------------------------------------------------------
> Report of Referee A -- LF15118/Masui
> ----------------------------------------------------------------------
> 
> In this paper, the authors consider the use the dispersion measure of
> radio transients as a proxy for radial distance, and compute the
> expected angular cross-power spectra of biased sources. With simple
> estimates of shot noise, the authors then argue that such power
> spectra should be observable, and suggest that this could be a
> powerful probe of large scale structure.
> 
> The calculation that the authors perform is relatively straightforward
> (though worthwhile) and uncontroversial. To be fair, there are a
> number of details about fast radio bursts (such as their source
> environments) that are swept under the rug. I agree with the authors,
> however, that such issues are beyond the scope of this short paper,
> since investigations into the nature of fast radio bursts is still an
> active area of research for the community. I would be happy to
> recommend publication once the authors explore the following points a
> little more fully:
> 
> i) The authors pick b_f = 1.3 to be the bias of the FRB sources. Was
> this arbitrary?

This was chosen to be roughly the value for star forming galaxies at these
redshifts as measured by the WiggleZ survey. The assumption here is that FRBs
might trace star formation, although this is entirely speculative. It also
provides some contrast with the electrons, since some terms are proportional to
b_e - b_f. This is now stated in the text.

> Also, it would be good to discuss briefly (perhaps
> with a new plot) the dependence of the paper’s conclusions on the
> value of b_f. In principle, variations of b_f could cause interesting
> effects, since b_f affects some of the terms but not others in Eq.
> (22). My guess is that the effects won’t be too big, since Fig. 2
> shows that the integral term is the dominant contribution, but it’d be
> nice to numerically confirm this. If anything, it would be helpful to
> know that the predictions are robust to uncertainties in b_f.

This was a very helpful comment and led us to include a b_f=0.7 line in Figure
3. We conclude that the sensitivity is indeed dominated by the integral term
and that a survey with 10000 events is unlikely to distinguish any signature of
the intrinsic clustering.

Time dependence of b_f should similarly be a small effect on the final
sensitivity. We've added a paragraph about the affect of the biases on the
various terms in the modelling section.

> 
> ii) To plot Fig. 3, the authors summed over pairs of radial bins.
> However, this hides some of the interesting statistical properties of
> the signal. For example, I would expect the cross power spectrum of
> two widely separated radial bins to be much lower than the cross power
> of two nearby bins. I would like to see some quantification of these
> radial correlations, as it would give readers a better sense for how
> the signal-to-noise scales with bin size, number of bins, etc.

Figure 2 is intended to give a hint of this structure. The integral term and
the cross term are fairly insensitive to radial separation since any structure
in the radial integral will correlate. The radial dependence is actually
dominated by A(\chi).  As shown in the figure, the local term drops rapidly at
higher separations. We would very much like to show more curves in the style of
Figure 2 but space constrains preclude this.

We've added a brief statement about this in the discussion.  We hope that the
reviewer finds this satisfactory.

> 
> Finally, just an extremely minor pet peeve of mine: can we replace the
> error bars in Fig. 3 with a band around the curve? Real data would
> have scatter, and the error bars wouldn’t be centered on the “truth”.

Our error bars depend on \ell binning and thus a band is inappropriate.
We have switched to filled boxes.  This is consistent with the style in 
Scott Dodelson's text book.

> 
> Thank you to the authors for an interesting read!