### 2015-08-05 writeup/summary.tex
started summary.tex

\documentclass{article}

\title{Mysterious radio flashes to be used to trace Universe's largest
structures}

\date{\today}

\author{Kiyoshi Wesley Masui and Kris Sigurdson}

\begin{document}

\maketitle

One of the most challenging aspects of astronomy is measuring distances.

\end{document}